---
title: "W271 Lab 3"
author: "Tiffany Jaya, Joanna Huang, Shan He, Robert Deng"
output: 
  pdf_document:
  toc: true
  number_sections: true
fontsize: 11pt
geometry: margin=1in
---

```{r}
# add packages
library(knitr)
# prevent source code from running off the page
opts_chunk$set(tidy.opts=list(width.cutoff=70), tidy=TRUE)
# remove all objects from current workspace
rm(list = ls())
# set seed number to reproduce results
set.seed(1)
# load data
# 1. ECOMPCTNSA: E-commerce retail sales as a percent of total sales
# https://fred.stlouisfed.org/series/ECOMPCTNSA
raw.sales <- read.csv('./data/ECOMPCTNSA.csv', header=TRUE, sep=',')
```

# Question 1: Forecasting using a SARIMA model

TODO: change the % and ARIMA model

Since the emergence of the internet, more and more people are shopping at online retailers than brick-and-mortar stores. E-commerce is on the rise, and we would like to see what percentage of total retail sales e-commerce is accounted for in the fourth quarter of 2017. With data from the US Census Bureau ranging from 1999 to 2016, we were able to estimate using the seasonal autoregresive integrated moving average model (or SARIMA for short) that e-commerce makes up approximately 10.44% of all retail sales. While the number does not seem substantial compare to the perceived value of e-commerce, we have to remember that retail sales include motor vehicles, gas stations, and grocery stores where e-commerce has yet to play a major role in the field. 

The SARIMA model that we use for the projected forecast is $\text{ARIMA}(0,0,0)(0,0,0)_4$. 

## Exploration Data Analysis 

The first step once we obtained the dataset was to examine its structure. 

```{r}
dim(raw.sales)
kable(summary(raw.sales)) 
sales <- ts(raw.sales$ECOMPCTNSA, start=c(1999,4), frequency=4)
head(sales); tail(sales)
```

We were able to determine that there was no missing value among the 69 observations with sales appearing to increase  overtime from 0.7% in the 4th quarter of 1999 to 9.5% in the 4th quarter of 2016. To confirm, we plot the time series as well as its associating MA(4) model. If the data expressed seasonality every quarter, the MA(4) model smooths out the variances and acts as an annual trend with the seasonal effects within each quarter removed.

```{r fig.show='hold', fig.align='center', out.width='49%'}
plot(sales, ylab = 'e-commerce sales (%)', main='E-commerce Quarterly Series')
lines(ma(sales, order=4, centre=T), col='blue')
acf(sales)
```

Given the upward trend and increasing variance, the series is non-stationarity with quarterly seasonality. The autocorrelation function further substantiates the series's non-stationarity because of its slow decay and multiple statistically significant values that are large and positive. For this reason, we will perform two operations. One, we will difference the series to stabilize the mean. And two, we will apply logarithmic transformation to stabilize the variance. We verified whether or not differencing is necessary using the unit root test.

```{r}
# unit root test
# H0: stationary, HA: p < 0.05, non-stationary
kpss.test(sales)
# number of differences required
nsdiffs(sales)
```

Basdfa on Kwiatkowski-Phillips-Schmidt-Shin (KPSS) test. 
We differenced the data twice in order to obtain a stationary series.

```{r fig.show='hold', fig.align='center', out.width='49%'}
# first-order differenced series plot
plot(diff(sales, difference=1))
# first-order differenced log-transformed series plot
plot(diff(log(sales), difference=1))
# first-order differenced log-transformed series
tsdisplay(diff(log(sales), lag=4, difference=1))
# second-order differenced log-transformed series
tsdisplay(diff(log(sales), lag=4, difference=2))
```
Another popular unit root test is the Kwiatkowski-Phillips-Schmidt-Shin (KPSS) test. This reverses the hypotheses, so the null-hypothesis is that the data are stationary. In this case, small p-values (e.g., less than 0.05) suggest that differencing is required.

nsdiffs() which uses seasonal unit root tests to determine the appropriate number of seasonal differences required.

```{r}
nsdiffs(log(sales))
```



 and not to mention its multiple statistically significant values as a result

stochastic trend: ARIMA(0,1,0)(0,0,1)_4
seasonal terms contain stochastic trend: ARIMA(0,0,0)(0,1,1)_4




exponential decay at lag 4, 8, 12 
exponential decay in the seasonal lags of the ACF
a significant spike at lag 4 in PACF

A significant value occurs at lag 1, suggesting that a more complex model may b needed 


In the plots of the seasonally differenced data, there are spikes in the PACF at lags 4, but nothing at seasonal lags in the ACF. This may be suggestive of a seasonal AR(1) term.




```{r}
# lag 1 autocorrelation
acf(sales.quarter)$acf[2]
```
ACF shows a significant autocorrelation at lag 4 might indicate that seasonal adjustment is not adequate






```{r}
# summary of the values for each season
boxplot(sales.quarter ~ cycle(sales.quarter))
```

Looking at the quarterly series graph, we can see that the data are non-stationarity with some seasonal trend. So we take the first order difference. 



```{r}

```

```{r}

```



The increasing trend is no longer apparent in the plots of the differenced series

First-order differencing successfully removed the increasing trend in the series.

```{r}
tsdisplay(diff(log(sales.quarter), 4))
```


```{r}
# remove both linear trend and additive seasonal effects in a quarterly series
tsdisplay(diff(diff(log(sales.quarter), 4)))
```

d = 1 because first-order differencing successfully removed the trend in the series

(0,0,0)(0,1,1)

AR, difference, MA
ARIMA (non-seasonal)(seasonal)
ARIMA (0,1,0)(0,1,2)


Our aim now is to find an appropriate ARIMA model based on the ACF and PACF.

the significant spike at lag 4 in the ACF suggests a seasonal MA(1) component and no other significant spikes
the PACF shows an exponential decay in seasonal lags, that is at lags 4, 8, 12...


```{r}
m1 <- Arima(sales.quarter, order=c(0,1,0), seasonal=c(0,1,1))
tsdisplay(residuals(m1))
AIC(m1)
```

```{r}
m2 <- Arima(sales.quarter, order=c(0,1,0), seasonal=c(0,1,2))
tsdisplay(residuals(m2))
AIC(m2)
```
```{r}
m3 <- Arima(sales.quarter, order=c(0,1,0), seasonal=c(0,1,3))
tsdisplay(residuals(m3))
AIC(m3)
```

```{r}
m4 <- Arima(sales.quarter, order=c(0,1,0), seasonal=c(0,1,4))
tsdisplay(residuals(m4))
AIC(m4)
```

```{r}
m5 <- Arima(sales.quarter, order=c(0,1,0), seasonal=c(0,1,5))
tsdisplay(residuals(m5))
AIC(m5)
```

# Question 2: Learning how to use the xts library

If we could select one company to represent the e-commerce trend, Amazon is it. 