---
title: "Tiffany's Work"
author: "Tiffany Jaya"
date: "7/24/2018"
output: html_document
---

```{r}
# add packages
library(car)
library(dplyr)
library(astsa)
library(Hmisc)
library(forecast)
library(ggplot2)
library(fpp)
library(xts)
library(MASS)
# prevent source code from running off the page
opts_chunk$set(tidy.opts=list(width.cutoff=70), tidy=TRUE)
# remove all objects from current workspace
rm(list = ls())
# set seed number to reproduce results
set.seed(1)
# load data
# 1. ECOMPCTNSA: E-commerce retail sales as a percent of total sales
# https://fred.stlouisfed.org/series/ECOMPCTNSA
sales <- read.csv('./data/ECOMPCTNSA.csv', header=TRUE, sep=',')
# 2. AMAZ.csv: Amazon stock sales
amazon <- read.csv('./data/AMAZ.csv', header=TRUE, sep=',')
# 3. UMCSENT.csv: University of Michigan Consumer Sentiment Index
# https://fred.stlouisfed.org/series/UMCSENT/ 
sentiment <- read.csv('./data/UMCSENT.csv', header=TRUE, sep=',')
# 4. bls_unemployment.csv: Bureau of Labor Statistic unemployment rate
# https://data.bls.gov/timeseries/LNS14000000 
unemp <- read.csv('./data/bls_unemployment.csv', header=TRUE, sep=',')
```


# Introduction

* summarize the question being asked
* result 

Data: E-Commerce retail sales as a percent of total sales

# EDA

* anomalies
* missing values
* potential top/bottom code

## Univariate Analysis

### ECOMPCTNSA

*ECOMPCTNSA.csv*, contains quarterly data of E-Commerce Retail Sales as a Percent of Total Sales. The data can be found at: https://fred.stlouisfed.org/series/ECOMPCTNSA.


```{r}
sales
```

--------------------------------------------------------------------------------

Source: U.S. Bureau of the Census   Release: Quarterly Retail E-Commerce Sales  
Units:  Percent, Not Seasonally Adjusted

E-commerce retail sales as a percent of total sales

Frequency:  Quarterly

E-commerce sales are sales of goods and services where the buyer places an order, or the price and terms of the sale are negotiated over an Internet, mobile device (M-commerce), extranet, Electronic Data Interchange (EDI) network, electronic mail, or other comparable online system. Payment may or may not be made online.

--------------------------------------------------------------------------------

Build a Seasonal ARIMA model and generate quarterly forecast for 2017. Make sure you use all the steps of building a univariate time series model between lecture 6 and 9, such as checking the raw data, conducting a thorough EDA, justifying all modeling decisions (including transformation), testing model assumptions, and clearly articulating why you chose your given model. Measure and discuss your model’s performance. Use both in-sample and out-of-sample model performance. When training your model, exclude the series from 2015 and 2016. For the out-of-sample forecast, measure your model’s performance in forecasting the quarterly E-Commerce retail sales in 2015 and 2016. Discuss the model performance. Also forecast beyond the observed time-period of the series. Specifically, generate quarterly forecast for 2017.

--------------------------------------------------------------------------------

1. Seasonal ARIMA model, quarterly forecast for 2017

```{r}
ts.sales <- ts(sales$ECOMPCTNSA, start=c(1999,4), frequency=4)
# exclude series from 2015 and 2016
ts.sales.train <- ts(ts.sales[time(ts.sales) < 2015], start=c(1999,4), frequency=4)
ts.sales.train <- ts(ts.sales[time(ts.sales) < 2015], start=c(1999,4), frequency=4)
```


2. Build a univariate time series model lecture 6 and lecture 9
2.a. check raw data

2.b. conduct thorough EDA

2.c. justify modeling decisions, including transformations

2.d. testing model assumptions

2.e. articulting why you chose your given model

3. Measure and discuss model's performance (in-sample and out-of-sample model performance), exclude series 2015-2016

3.a. in-sample, exclude-series 2015-2016

3.b. out-of-sample, in 2015 and 2016

4. Generate quarterly forecast for 2017


### AMAZ

```{r}
amazon
```

### UMCSENT

```{r}
sentiment
```

Units:  Index 1966:Q1=100, Not Seasonally Adjusted

Frequency:  Monthly

At the request of the source, the data is delayed by 1 month. To obtain historical data prior to January 1978, please see FRED data series UMCSENT1.

This data should be cited as follows: "Surveys of Consumers, University of Michigan, University of Michigan: Consumer Sentiment© [UMCSENT], retrieved from FRED, Federal Reserve Bank of St. Louis https://fred.stlouisfed.org/series/UMCSENT/, (Accessed on date)"

### bls_unemployment

```{r}
unemp
```



Series Id:           LNS14000000
Seasonally Adjusted
Series title:        (Seas) Unemployment Rate
Labor force status:  Unemployment rate
Type of data:        Percent or rate
Age:                 16 years and over

# Modeling 

* rational of decision for selecting the model

# Conclusion






